---
title: "Download and plot worldclim scenario data  tmax"
author: "Kami Makabe"
date: "5/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE, warning = FALSE)
```

### Load required libraries
If using r for the first time, you need to install the respective packages fisrt.
The package names as similar to the libraries.

```{r}
library(raster)
library(magrittr)
library(rnaturalearth)
library(dplyr)
library(reshape)
library(ggplot2)
library(ggeasy)
library(stringr)
```

### Prepare your data
We will need to manually download our scenario data from [worldclim](https://worldclim.org/data/cmip6/cmip6climate.html), and store it locally.
Alternatively, use direct download for example 'ssp126_40url<- "https://biogeo.ucdavis.edu/data/worldclim/v2.1/fut/2.5m/wc2.1_2.5m_prec_BCC-CSM2-MR_ssp126_2021-2040.zip"
download.file(ssp126_40url, destfile="wc2.1_2.5m_prec_BCC-CSM2-MR_ssp126_2021-2040.zip")
unzip("wc2.1_2.5m_prec_BCC-CSM2-MR_ssp126_2021-2040.zip")'

Set the path to your downloaded files
```{r}
path126<-"C:/Workspace/RMardown_NAPS/Open-NAPs-Data/opennaps/WC_Fut_data/tmax126"
path245<-"C:/Workspace/RMardown_NAPS/Open-NAPs-Data/opennaps/WC_Fut_data/tmax245"
path370<-"C:/Workspace/RMardown_NAPS/Open-NAPs-Data/opennaps/WC_Fut_data/tmax370"
path585<-"C:/Workspace/RMardown_NAPS/Open-NAPs-Data/opennaps/WC_Fut_data/tmax585"
```

### Import files into r
Then load the files into your r environment. 
As a tip, since we already know our data is in .tif format, we will give r the instructions to only load files with the .tif extension. This works well, incase you use the directory for other file types.

```{r}
tx126_tif_only<-list.files(path126,full.names = TRUE, pattern = ".tif$")
tx245_tif_only<-list.files(path245,full.names = TRUE, pattern = ".tif$")
tx370_tif_only<-list.files(path370,full.names = TRUE, pattern = ".tif$")
tx585_tif_only<-list.files(path585,full.names = TRUE, pattern = ".tif$")
```


Each of the downloads has multiple files in it and this step will pile them together into one layer, also known as a rasterstack

```{r}

ssp126_stack<-stack(tx126_tif_only)
ssp245_stack<-stack(tx245_tif_only)
ssp370_stack<-stack(tx370_tif_only)
ssp585_stack<-stack(tx585_tif_only)
```

### Get geom boundary 
Define your bounding geometry. 
This is where data to define your area of interest goes.
You may upload your own geometry,such as shapefile or use any of the mapservices available in r.Two options are shown below

```{r, echo=FALSE}

malawi<-rnaturalearth::ne_countries(country ='malawi', returnclass='sf') # option 1
#mwi<-getData('GADM', country='MWI', level=0) # option 2, remove '#' to execute
```


### Subset your data

Use the geometry in previous step to subset your data to your area of interest.
We will use two raster functions 'crop' and 'mask' for this task.

```{r}

ssp126malawi<-crop(ssp126_stack,extent(malawi)) ## here you can use your shapefile from previous step or get the bounding coordinates from any map service and enter them in the order (west, east, south, north)
ssp126malawim<-mask(ssp126malawi,malawi)
#plot(ssp245malawim)  # remove the '#' sign to plot the individual stacked .tifs
ssp245malawi<-crop(ssp245_stack,extent(malawi)) 
ssp245malawim<-mask(ssp245malawi,malawi)

ssp370malawi<-crop(ssp370_stack,extent(malawi)) 
ssp370malawim<-mask(ssp370malawi,malawi)

ssp585malawi<-crop(ssp585_stack,extent(malawi)) 
ssp585malawim<-mask(ssp585malawi,malawi)
```

### Convert subset raster to data frame
Convert the subset data into data frame format using the function 'as.data.frame' and the 'melt' function to convert our data frame from a wide table to a long table format


```{r}

ssp126df<-as.data.frame(ssp126malawim, xy=TRUE, na.rm=TRUE, stringAsFactors=TRUE)%>%
  melt(id.vars=c('x','y'))
ssp245df<-as.data.frame(ssp245malawim, xy=TRUE, na.rm=TRUE, stringAsFactors=TRUE)%>%
  melt(id.vars=c('x','y'))
ssp370df<-as.data.frame(ssp370malawim, xy=TRUE, na.rm=TRUE, stringAsFactors=TRUE)%>%
  melt(id.vars=c('x','y'))
ssp585df<-as.data.frame(ssp585malawim, xy=TRUE, na.rm=TRUE, stringAsFactors=TRUE)%>%
  melt(id.vars=c('x','y'))

```


### Plot  all the rasters from the df and the bounding geometry

```{r}

ggplot()+
  geom_raster(data=ssp126df, aes(x=x, y=y, fill=value))+
  facet_wrap(~variable)+
  geom_sf(fill='transparent', data = malawi)+
  scale_fill_gradient(low = 'yellow', high = 'red', name='째C')+
  xlab('Longitude')+ylab('Latitude')+ ggtitle('Monthly Maximum Temperature')+
  theme_bw()
  
```


### Select and plot single month

```{r}

xmonthdf<-as.data.frame(ssp126malawim[[06]], xy=TRUE, stringsAsFactors=FALSE)%>%
  melt(id.vars=c('x','y'))
ggplot()+
  geom_raster(data=xmonthdf, aes(x=x, y=y, fill=value))+
  xlab('Longitude')+ylab('Latitude')+
  geom_sf(fill='transparent', data = malawi)+
  scale_fill_gradient(low = 'yellow', high = 'red', name='째C')+
  ggtitle('Monthly Max Temperature-June')+
  theme_bw()
  
```

### Add column 'Trend' to dataframes
The next step would be to combine the 4 dataframes into one so that we can create plots easily.
But since the data variables are identical, with no column to distinguish them, we will add a column called 'Trend' to each and populate it with respective values.
Then combine the 4 dataframes using the base function 'rbind'.

```{r}
Trend<-'ssp126'
ssp126df<-cbind(ssp126df,Trend)
Trend<-'ssp245'
ssp245df<-cbind(ssp245df,Trend)
Trend<-'ssp370'
ssp370df<-cbind(ssp370df,Trend)
Trend<-'ssp585'
ssp585df<-cbind(ssp585df,Trend)

tmax<-rbind(ssp126df,ssp245df,ssp370df,ssp585df)

```

### Add year and month columns to the data frame
Next, extract month and year values from the variable name and add them as separate columns

```{r}

Year<-substr(tmax$variable,36,44)
Month<-as.numeric(substr(tmax$variable,46,47))
tmax<-cbind(tmax, Month, Year)
```
Rename column 'value' to 'tmax'

```{r}
colnames(tmax)[colnames(tmax)=="value"]<-"tmax"
```

### Plot monthly mean time series
1. To plot the monthly data, first we have to group the data by 'Trend' and 'Month'
2. Then calculate monthly mean from the tmax values and feed this data into a plot; we use qplot from the package 'ggeasy'

```{r}

txmon<-tmax%>%group_by(Trend, Month)%>%
  summarise(across(contains("tmax"), ~mean(tmax, na.rm=TRUE)))

txmon$Month<-factor(txmon$Month,levels = c('1','2','3','4','5','6','7','8','9','10','11','12'))

qplot(factor(Month),tmax, data = txmon, geom = "line", group = Trend, color= Trend) +
     #facet_grid(Trend ~ ., scale = "free") +
     scale_x_discrete() +
  ggtitle('Maximum Temperature Trend - Monthly')+
     easy_rotate_x_labels(angle = 90, side = "right")+
  ylab('Max. Temp (째C)')+ xlab('Month')
```

### Plot annual mean time series

1. To plot the yearly data, first we have to group the data by 'Trend' and 'Year'
2. Then calculate yearly mean from the tmax values and feed this data into a plot; we use qplot from the package 'ggeasy'

```{r}

txann<-tmax%>% group_by(Year, Trend)%>% 
  summarise(across(contains("tmax"), ~mean(tmax)))
tstxann<- ts(txann$tmax,start = 2021,end=2100, frequency = 1)
#plot(tstxann)

qplot(Year,tmax, data = txann, geom = "line", group = Trend, color= Trend) +
     #facet_grid(Trend ~ ., scale = "free") +
     scale_x_discrete() +
  ggtitle('Maximum Temperature Trend - Annual')+
     easy_rotate_x_labels(angle = 90, side = "right")+
  ylab('Max. Temp (째C)')+ xlab('Period (Years)')
#plotly::ggplotly(myplot)
```

